---
title: "VIS_assignment 01"
author: "Mengyao"
date: "2020/9/11"
output:   
  html_document:
    theme: lumen
    toc: true
    toc_float: true

---

# Load packages
```{r, message=FALSE}
library(tidyverse)
library(ggplot2)
library(knitr)
```

# Load data
As I had connection issue to get data directly from census, I tried to download three tables from the website and combined them.
I downloaded 1 zippd file named Means of transportation to work by class of workers.zip from data.census.gov. The zipped file contains three files:

1. The dataset (Means of transportation to work by class of workers_data.csv)
2. A list of variable names and descriptions (Means of transportation to work by class of workers_var.csv)
3. A description of the data source (Means of transportation to work by class of workers_data.txt)

```{r, results='hide'}
unzip("Means of transportation to work by class of workers.zip")
tran_variables <- read.csv("Means of transportation to work by class of workers_var.csv")
tran_data <- read.csv("Means of transportation to work by class of workers_data.csv", skip =1)
colnames(tran_data)
```

# Choosing variables
The table is based on 5-year American Community Survey of tracts in Middlesex County, Massachusetts in 2018.
I want these variables from the table:
1. The population of the tract (Continuous)
2. The percentage of people who use car to commute (Continuous)
3. The percentage of people who walk or use public transportation to commute (Continuous)
```{r, results='hide'}

data_commute <- tran_data %>%
  mutate(population =`Estimate..Total`,
         pct_car = `Estimate..Total..Car..truck..or.van...drove.alone`/`Estimate..Total`,
         pct_pub_trans = `Estimate..Total..Public.transportation..excluding.taxicab.`/`Estimate..Total`)%>%
  mutate(geoid = substr(id, 10, 20)) %>%
  select(geoid, population, pct_car, pct_pub_trans)
data_commute
  
```

# Load the 2nd dataset about income

```{r, results='hide'}
unzip("Income.zip")
income_variables <- read.csv("income_vars.csv")
income_data <- read.csv("income_data.csv", skip =1)
colnames(tran_data)
```
The table is based on 5-year American Community Survey of tracts in Middlesex County, Massachusetts in 2018.
I want these variables from the table:
1. Whether foreign residents are richer (Categorical)
2. The median income (Continuous)
```{r, results='hide'}
data_income <- income_data %>%
  mutate(median_income = `Estimate..Median.income.in.the.past.12.months.....Total`,
         rich_foreign = `Estimate..Median.income.in.the.past.12.months.....Total..Foreign.born` > `Estimate..Median.income.in.the.past.12.months.....Total..Born.in.state.of.residence`)%>%
  mutate(geoid = substr(id, 10, 20)) %>%
  select(geoid, median_income, rich_foreign)
data_income
```

# Load the 3rd dataset about race

```{r, results='hide'}
unzip("Race.zip")
race_variables <- read.csv("race_vars.csv")
race_data <- read.csv("race_data.csv", skip =1)
colnames(race_data)
```
The table is based on 5-year American Community Survey of tracts in Middlesex County, Massachusetts in 2018.
I want this variable from the table:
1. The majority race of the tract (Categorical)
```{r, results='hide'}
data_race <- race_data %>%
  mutate(majority_race = case_when(
    `Estimate..Total..White.alone`/`Estimate..Total` > 0.5 ~ "white",
    `Estimate..Total..Black.or.African.American.alone`/`Estimate..Total` > 0.5 ~ "black",
    `Estimate..Total..Asian.alone`/`Estimate..Total` >0.5 ~ "asian",
    (`Estimate..Total..White.alone`/`Estimate..Total`+`Estimate..Total..Black.or.African.American.alone`+`Estimate..Total..Asian.alone`+`Estimate..Total..White.alone`)/`Estimate..Total` < 0.5 ~ "other", TRUE ~ "None"))%>%
  mutate(geoid = substr(id, 10, 20)) %>%
  select(geoid, majority_race)
data_race
```

# Join data
This dataset includes 318 tracts. The first ten rows are presented here as a formatted table.
```{r}
all_data_01 <- inner_join(data_commute, data_income, by="geoid")
all_data <- inner_join(all_data_01, data_race, by="geoid")
knitr::kable(all_data[1:10,])
```
```{r}
unique(all_data$majority_race)
```

# Plot 1
Plotting a majority of my variables
```{r}
ggplot(all_data,
       aes(x= population, 
           y= pct_car, 
           size = pct_pub_trans, 
           color = majority_race,
           shape = rich_foreign)) +
  geom_point() +
  scale_x_continuous(name = "Tract population") + 
  scale_y_continuous(name = "Percentage of car commuter",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100,by=10), "%", sep=""))+
  scale_size_continuous(name ="Percentage of public transportation commuter") +
  scale_color_discrete(name = "Majority race",
                       labels = c("No Majority", 
                                  "Asian", 
                                  "White"))+
  scale_shape(name = element_blank(),
              labels = c("Foreign residents earn more", "Foreign residents earn less")) +
  theme_light() +
  theme(legend.key.size = unit(0.1, "cm"))

```

# Plot 2
Percentage of people using car for commute and median income
```{r}
ggplot(all_data,
       aes(x=pct_car,
           y=median_income)) +
  geom_point() +
  scale_x_continuous(name = "Percentage of people using car for commuting",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100, by=10), "%", sep = ""))+
  scale_y_discrete(name = "Median income",
                   breaks = seq(10000, 110000, by=10000)) +
  theme_light()
```

# Plot 3
Majority race and percentage of people using public transportation for commute
```{r}
ggplot(all_data, 
       aes(x=majority_race,
           y=pct_pub_trans,
           fill= majority_race)) +
  geom_dotplot(binaxis = "y", dotsize = 0.5) +
  scale_x_discrete(name ="Majority race",
                   labels = c("No Majority", "Asian", "White")) +
  scale_y_continuous(name = "Percentage of people using public transportation for commute",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100,by=10), "%", sep="")) +
  labs(fill = "Majority race") +
  scale_fill_discrete(name = "Majority race",
                      breaks = c("No Majority", "Asian", "White")) +
  theme_minimal()
```

# Plot 4
Heat map of richer foreigners, majority race and percentage of car commuters
```{r}
ggplot(all_data,
       aes(x = majority_race,
           y = rich_foreign,
           fill = pct_car)) +
  geom_tile() +
  scale_x_discrete(name ="Majority race",
                   labels = c("No Majority", "Asian", "White")) +
  scale_y_discrete(name = "Foreigners are richer",
                   labels = c("True", "False")) +
  scale_fill_distiller(palette = "RdPu") +
  labs(fill = "Percentage of car commuters") +
  theme_classic()
```

# Plot 5
Scatter plot of percentage of car commuters and public transportation commuters
```{r}
ggplot(all_data,
       aes(x = pct_car,
           y = pct_pub_trans,
           size = population)) +
  geom_point (alpha = 0.5, shape = 21, color = "black") +
  scale_x_continuous(name = "Percentage of people using car for commuting",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100, by=10), "%", sep = ""))+
  scale_y_continuous(name = "Percentage of people using public transportation for commute",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100,by=10), "%", sep="")) +
  theme_minimal()
```

# Plot 6
Violin Plot of percentage of public transportation user and majority race
```{r}
ggplot(all_data,
       aes(x = majority_race,
           y = pct_pub_trans,
           fill = majority_race)) +
  geom_violin(width = 2.1, size = 0.2, alpha = 0.75) +
  scale_x_discrete(name ="Majority race",
                   labels = c("No Majority", "Asian", "White")) +
  scale_y_continuous(name = "Percentage of people using public transportation for commute",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100,by=10), "%", sep="")) +
  theme_minimal()
```

# Plot 7
Box plot of percentage of public transportation user and whether foreigner is richer
```{r}
ggplot(all_data,
      aes(x = rich_foreign,
          y = pct_pub_trans,
          fill = rich_foreign)) +
  geom_boxplot() + 
  scale_x_discrete(name = "Foreigners are richer",
                   labels = c("True", "False")) +
  scale_y_continuous(name = "Percentage of people using public transportation for commute",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100,by=10), "%", sep="")) +
  geom_jitter(color = "blue", size = 0.4, alpha = 0.9) +
  theme_light()
```

# Plot 8
Area chart of percentage of residents that use car to commute
```{r}
ggplot(all_data,
       aes(x = pct_car)) +
  geom_density(fill = "#b31049", alpha = 0.8) +
  scale_x_continuous(name = "Percentage of people using car for commuting",
                     breaks = seq(0,1,by=0.1),
                     labels = paste(seq(0,100, by=10), "%", sep = ""))+
  theme_light()
```

# Plot 9
Majority race and median income
```{r}
ggplot(all_data,
       aes(x = median_income,
           y = majority_race)) +
  geom_point() +
  geom_segment(aes(x=0, 
                   xend = median_income, 
                   y = majority_race, 
                   yend = majority_race)) +
  scale_x_discrete(name = "Median income",
                     breaks = seq(10000, 110000, by=10000)) +
  scale_y_discrete(name ="Majority race",
                   labels = c("No Majority", "Asian", "White")) +
  theme_minimal()
```

# Plot 10
Stacked bar chart of majority race and population
```{r}
ggplot(all_data,
       aes(x = majority_race,
           y = population,
           fill = majority_race)) +
  geom_bar(stat = "identity", width = 0.5) +
  scale_x_discrete(name ="Majority race",
                   labels = c("No Majority", "Asian", "White")) +
  scale_y_continuous(name = "Population") +
  theme_bw()
```








